---
title: "05 MAST DE testing"
output: html_notebook
---

#DE Testing for the identification of cell type-specific transcription factor signatures

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=8, fig.path='Figs/', echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, warning=FALSE}
# Load required packages
library(DropletUtils)
library(MAST)
library(BiocSingular)
n_cores <- 4
```

```{r}
obj <- readRDS("")
```

Loading of transcription factor (TF) reference list 
```{r}
tfs <- read.table("", header = T, sep = "\t")
tfs <- tfs$Symbol
```

Remove genes that are not TFs, and any TFs that are not detected in any cell
```{r}
mast_obj <- SceToSingleCellAssay(obj)
mast_obj <- mast_obj[rownames(mast_obj) %in% tfs, ]
mast_obj <- mast_obj[rowSums(assay(mast_obj)) > 0, ]
```

Create a binary condition for the 1-vs-rest DE testing across clusters
```{r}
group <- "Sst"
group_number <- "7"
clusters <- colData(mast_obj)$cluster_redone_v2
clusters <- as.numeric(as.character(clusters))
clusters[clusters == group_number] <- group
clusters[clusters != group] <- "Other"
colData(mast_obj)$Condition <- as.factor(clusters)
```

```{r}
colData(mast_obj)$detected <- scale(colData(mast_obj)$detected)
```

```{r}
cond <- factor(colData(mast_obj)$Condition)
cond <- relevel(cond, "Other")
colData(mast_obj)$Condition <- cond
```

Data fitting
```{r}
options(mc.cores = n_cores)
zlm_cond <- zlm(
  formula = ~Condition + batches + detected,
  sca = mast_obj,
  parallel = TRUE,
)
```

Likelihood test
```{r}
summary_cond <- summary(zlm_cond, doLRT = paste0("Condition", group))
```

Data collection
```{r}
# Creates data table
summary_dt <- summary_cond$datatable
results <- merge(
  summary_dt[
    contrast == paste0("Condition", group) & component == "H",
    .(primerid, `Pr(>Chisq)`)
  ],
  summary_dt[
    contrast == paste0("Condition", group) & component == "logFC",
    .(primerid, coef)
  ],
  by = "primerid"
)

#Drop where coef cannot be calculated
results <- results[!is.na(results$coef)]
```

Multiple test correction and FDR-based significance threshold
```{r}
results[, fdr := p.adjust(`Pr(>Chisq)`, "fdr")]
dexp <- results[results$fdr < 0.05, , drop = F]
dexp <- dexp[order(dexp$fdr), ]
```

Output column names:
PrimerId = gene name 
Pr(>Chisq) = raw probability 
Coef = log2FC
fdr = FDR
